---
title: "Exercise 8.2: Assignment07"
author: "Katie Adams"
date: "25/04/2021"
output: pdf_document
---

```{r, include=TRUE}
library(tinytex)
knitr::opts_chunk$set(echo = TRUE)

## Load the `data/r4ds/heights.csv` and assign a variable
heights_df <- read.csv("C:/Users/kadams/OneDrive - Suncor Energy Inc/DSC Masters Program/GitHub/DSC520/Assignments/dsc520/data/r4ds/heights.csv", header=TRUE)
```

## Assignment 07

1. Fit a linear model
```{r, include=TRUE}
earn_lm <-  lm(earn ~ ., data = heights_df)
earn_lm
```
2. View the summary of your model
```{r, include=TRUE}
summary(earn_lm)
```
3. Compute deviation 
```{r, include=TRUE}
mean_earn <- mean(heights_df$earn)
mean_earn
```
4. SST: Corrected Sum of Squares Total
```{r, include=TRUE}
sst <- sum(anova(earn_lm)[1:6,2])
```
6. Residuals
```{r, include=TRUE}
residuals <- anova(earn_lm)[6,]
```
7. SSE: Sum of Squares for Error
```{r, include=TRUE}
sse <- anova(earn_lm)[6,2]
```
6. SSM: Corrected Sum of Squares for Model
```{r, include=TRUE}
ssm <- sst-sse #changed equation for easier running of different equations
```
8. R Squared
```{r, include=TRUE}
r_squared <- ssm/sst
```
9. Number of observations
```{r, include=TRUE}
n <- nrow(heights_df)
```
10. Number of regression paramaters
```{r, include=TRUE}
p <- 8
```
11. Corrected Degrees of Freedom for Model
```{r, include=TRUE}
dfm <- p-1
```
12. Degrees of Freedom for Error
```{r, include=TRUE}
dfe <- n-p
```
13. Corrected Degrees of Freedom Total:   DFT = n - 1
```{r, include=TRUE}
dft <- n-1
```
14. Mean of Squares for Model:   MSM = SSM / DFM
```{r, include=TRUE}

msm_age <- anova(earn_lm)[4,3]
msm_height <- anova(earn_lm)[1,3]
msm_sex <- anova(earn_lm)[2,3]
msm_ed <- anova(earn_lm)[3,3]
msm_race <- anova(earn_lm)[5,3]
```
15. Mean of Squares for Error:   MSE = SSE / DFE
```{r, include=TRUE}
mse <- anova(earn_lm)[6,3]
```
16. Mean of Squares Total:   MST = SST / DFT
```{r, include=TRUE}
mst <- sum(anova(earn_lm)[1:6,3])
```
17. F Statistic
```{r, include=TRUE}
f_score_age <- anova(earn_lm)[4,4]
f_score_height <- anova(earn_lm)[1,4]
f_score_sex <- anova(earn_lm)[2,4]
f_score_ed <- anova(earn_lm)[3,4]
f_score_race <- anova(earn_lm)[5,4]
```
18. Adjusted R Squared R2 = 1 - (1 - R2)(n - 1) / (n - p)
```{r, include=TRUE}
first_earn <- 1-r_squared
second_earn <- dft
third_earn <- dfe

adjusted_r_squared <- 1 - ((first_earn*second_earn)/third_earn)
```